Transition path sampling based calculations of free energies for enzymatic reactions: the case of human methionine adenosyl transferase and plasmodium vivax adenosine deaminase

### 2022-06-09 tex/resubmission1/response/response.tex

%\begin{response}
%{The current work shows the application of the BOLAS method to two enzyme-catalyzed reactions. 
%The BOLAS method is a transition path sampling algorithm that can be extended to calculate free 
%energies along a pre-defined reaction coordinate. There are already several methods to calculate 
%free energy profiles and many have been applied to enzymatic reactions, but none is a black-box, 
%and therefore the implementation and application of this method to an enzymatic reaction is an 
%important advance. The text is well written and clear.}
%reply
%TPS based free energy calculations have been attempted previously for the calculations
%involving model problems involving analytical potential energy surfaces or for 
%configurational or structural changes of an enzyme. A chemical reaction catalyzed by 
%an enzyme can involve a complicated reaction coordinate and TPS based method using the 
%shooting algorithm provides a bias free technique for calculating the free energy
%of the reaction.

%\end{response}

\begin{response}
{First one needs to know how the specific parameters of the method affect 
the results. Or at least, how or why they were chosen. If not for all parameters, 
at least for the most relevant ones (but this needs also to be specified). 
For example:
* why use 150 reactive trajectories?}
%reply

%The application of TPS for stuyding the kinetics and transition states of 
%enzyme catalyzed reactions has been repeorted several times by our group as well
%as by others we decided to not list out every detail. But we have now included 
%some details about the choice the number of trajectories and so on.
The number of reactive trajectories in a TPS ensemble is decided based 
on how quickly the shooting algorithm results in decorrelation of the reactive 
trajectories from the initial biased trajectory. If the decorrelation is fast 
then fewer trajectoreis are sufficient for further committor analysis. 
To be sure that we do not use trajecotries that are correlated to the initial 
biased trajectory we ignore the first 100 trajectories for committor analysis. 
Additionally we have added a discussion on the acceptance percentage of reactive 
trajectories in the manuscript. 
\end{response}

@@

\begin{response}
  {why should each reaction last 0.5 ps? How is it known if that time is too short 
  or too long?}
We used longer trajectories (1 ps) and checked to make sure that the reaction occurs within
the time that we used (0.5 ps) before collecting the TPS ensemble.

\end{response}

{\begin{response}
{* why 20 windows?}
%reply
This choice was made based on the order parameter window that we are scanning.
20 windows results in a window size of 0.39 angstroms and we use a overlap of
0.09 angstroms on either side of the window to make sure that we have configurations
that are common to neighboring windows which is necessary to obtain a continuous free energy
profile. Smaller window sizes result in inefficient sampling because the TPS based method
does not include bias forces to keep the trajectories within the windows and we end up
sampling outside the windows especially in regions where the free energy has a steeper slope.  
\end{response}

\begin{response}
{* When calculating the commitor, you cannot say "repeating this procedure many times". 
How many are needed to obtain a given precision or to obtain the given figure?
\newline
* As before, when calculating the separatrix, you can not say "repeating this 
process for several trajectories".}
%reply
Since this is the discussion of the committor analysis and committor distribution 
analysis in general and not for the specific case for the enzymes we had not
provided the number of trajectories that were used in these calculations.
\end{response}

\begin{response}{Some of these questions arise because nowhere in 
the manuscript are confidence intervals given. Neither in the energy 
profile nor in the transition state structure distances (Fig. 5). 
Therefore the effect of these parameters on the precision of the 
results is not known.
}
We have added a supporting information file which contains analysis on the transition
state ensemble providing more information about the structures that we had calculated. 
The transition state structure depicted in Fig. 5 is a representative
from the ensemble of transition states.
\end{response}

@@

% REPLY

\end{response}

\begin{response}{Beyond these general issues raised so far, there are some other points 
that deserve some attention.
\newline
1. How is the initial reactive trajectory generated? Although obtaining an initial reactive 
trajectory is not the crux of the algorithm it is a non-trivial step that should be more 
detailed. In particular, for the pvADA enzyme, which involves a two-step reaction, how 
do the way these 2 steps take place in the initial trajectory condition of the TPS ensemble?}

% REPLY
The reaction catalyzed by the \textit{pv}ADA enzyme consists of 
the nucleophilic attack of the OH$^{-}$ on the C6 atom, proton transfer from the 
GLU 223 residue to the N1 atom in the adenosyl moeity and finally the 
the transfer of the proton from the OH$^{-}$ to the N6 atom resulting in the formation 
of NH$_3$ and inosine. We have included the time series of the key distances in 
the supporting information which shows that we collect trajectories which contain all these
steps taking the system from the reactant state to the product state with the order of events
as predicted from experiments. 
\end{response}

### 2022-06-09 tex/resubmission1/manuscript.tex
added a supporting information and writing some sections there

systems.

>>>>

systems. The percentage of TPS trajectories that were accepted were 24$\%$ and 
19$\%$ for the MAT2A and $pv$ADA systems, respectively.
For optimal sampling with a good balance between the number of 
accepted reactive trajectories and decorrelation from the initial constrained
trajectory as well from one another, the acceptance percentage 
is typically $\sim$ 40. \cite{dellago02AdvChemPhys123}  
For complex systems such as enzymes we find that the lower the acceptance 
percentage the quicker is the decorrelation. 
%While a 40$\%$ acceptance has been found to provide the best compromise between the
%number of accepted trajectories and decorrelation of the
%trajectories from one another, 
%we find that a lower acceptance
%ratio provides quicker decorrelation from the initial constrained
%trajectory, especially in more complex systems such as enzymes.

### 2022-06-09 tex/resubmission1/supp-info/si.tex
added a supporting information and writing some sections there

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage[scaled=0.92]{helvet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{tocentry}
\section{Initial reactive trajectory}
\begin{figure}
\includegraphics[scale=1]{figures/ada-oh-nh60.pdf}
\caption{The distance between the N6-H1 atoms and O1-H1 atoms as a function of time. The proton transfer from the hydroxide
anion to the NH$_2$ group occurs after the nucleophilic attack of the O1 atom on the C6 atom.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage{multirow}
\usepackage{tabularx,booktabs}
\usepackage{color, colortbl}
\usepackage{subcaption}
\usepackage[scaled=0.92]{helvet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{tocentry}
\section{Initial reactive trajectory}
The reaction catalyzed by \textit{pv}ADA is between adenosine and OH$^{-}$ anion resulting
in the formation of inosine. Experimental studies suggest that the GLU 223 residue 
protonates the N1 atom of the adenosyl moeity and this transfer of proton is almost complete 
at the transition state. \cite{Luo07JAmChemSoc129p8008} The protonation of the N1 atom 
as observed in a representative reactive trajectory from the TPS ensemble is shown in 
Fig. \ref{sfig:he2time}.

\begin{figure}
\includegraphics[scale=1]{figures/ada-ohe-nhe60.pdf}
\caption{The distance between the N1-HE2 atoms (d$_{\text{NH}}$) and the OE2-HE2 atoms 
(d$_{\text{OH}}$)in {\AA} as a function of time. The proton is 
transferred from the GLU 229 residue onto the N1 atom of the adenosyl moeity.}
\label{sfig:he2time}
\end{figure}

\begin{figure}
\includegraphics[scale=1]{figures/ada-oh-nh60.pdf}
\caption{The distance between the N6-H1 atoms and O1-H1 atoms as a function of time. The proton transfer from the hydroxide
anion to the NH$_2$ group occurs after the nucleophilic attack of the O1 atom on the C6 atom.}
\end{figure}

\section{Transition state ensemble for MAT2A enzyme}
\begin{table*}[ht!]
\caption{Order parameters d$_{\text{SC}}$ and d$_{\text{OC}}$ in {\AA} for the 11 transition states obtained from 
the TPS ensemble along with the mean and standard deviation (Stddev).}
\centering
\begin{tabular}{l c c}
\hline\hline
Transition state & d$_{\text{SC}}$ ({\AA}) & d$_{\text{OC}}$ ({\AA})\\
\hline
1 & 2.30 & 2.10 \\
2 & 2.31 & 2.08 \\
3 & 2.31 & 2.09 \\
4 & 2.34 & 2.14 \\
5 & 2.35 & 2.17 \\
6 & 2.41 & 2.16 \\
7 & 2.41 & 2.11 \\
8 & 2.40 & 2.15 \\
9 & 2.35 & 2.13 \\
10& 2.38 & 2.12 \\
11& 2.29 & 2.06 \\
\hline
Mean & 2.35 & 2.12 \\
Stddev & 0.04 & 0.03 \\
\hline\hline
\end{tabular}
%
\end{table*}

\section{Transition state ensemble for \textit{pv}ADA enzyme}
\begin{table*}[ht!]
\caption{Order parameters d$_{\text{OC}}$ and d$_{\text{NC}}$ in {\AA} for the 11 transition states obtained from 
the TPS ensemble along with the mean and standard deviation (Stddev).}
\centering
\begin{tabular}{l c c}
\hline\hline
Transition state & d$_{\text{OC}}$ ({\AA})& d$_{\text{NC}}$ ({\AA})\\
\hline
1& 1.48  &1.60 \\
2& 1.45  &1.51 \\
3& 1.48  &1.58 \\
4& 1.5   &1.57 \\
5& 1.55  &1.55 \\
6& 1.55  &1.64 \\
7& 1.52  &1.63 \\
8& 1.52  &1.68 \\
9& 1.49  &1.52 \\
10&1.50  &1.64 \\
11&1.49  &1.62 \\
\hline
Mean & 1.50 & 1.59 \\
Stddev & 0.03 & 0.05 \\
\hline\hline
\end{tabular}
%
\end{table*}

\bibliography{si}
